Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2024-12-23 main.tex
Sec. V-A-2 edited

    \paragraph{Task and edge weights}
    For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
    The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
    Each row corresponds to an execution of one task on one cluster node.
    Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
    to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
    size of files that the task sends to all its children.

    For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
    The same workflow can require different memory capacity and take different time to execute
    depending on the size of its input.
    We simulate these various runs by obtaining values corresponding to each input size.
    For each of the four families, there are five input sizes, so we run each workflow in five variants corresponding to these inputs

    Not all tasks have historical runtime data stored in the tables.
    In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
    For two more, around 40\% of tasks have no historical runtime data stored.
    Hence, in the absence of historical data about a task, we give it fixed weights.
    We give it an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
    These values align with the findings of~\cite{lotaru} about small tasks.
%
 %   \skug{insert more abt how the runtime system hanfles data reads}

@@

    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to $12$, thus having 72 processors in total.  % the whole cluster.
%
    Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
    ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and
    large memory (see Table~\ref{tab:procs}).

    \begin{table}[htb]
        \begin{center}
            \begin{tabular}{c|c|cc}
                \toprule
                Processor  %& Amount
                &  CPU speed   & \multicolumn{2}{c}{Memory size (GB)} \\
                name & (GHz) & {\em default} & {\em mem-constrained} \\
                \midrule
                local                    & 4                    & 16     & 1.6 \\
                A1                      & 32                   & 32     & 3.2 \\
                A2                      & 6                    & 64    & 6.4 \\
                N1                      & 12                   & 16     & 1.6 \\
                N2                      & 8                    & 8      & 0.8\\
                C2                      & 32                   & 192   &  19.2\\
                \bottomrule
            \end{tabular}
        \end{center}
        \caption{Cluster configuration.}
        \label{tab:procs}
    \end{table}

    We also consider a more constrained setting, by varying the cluster configuration. The
     {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
     but each node has 10 times less memory. Hence,  
     the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, 
     $A2$ has $6.4$~GB instead, of $64$~GB etc.
    The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).

>>>>

\paragraph{Task and edge weights}
For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
Each row corresponds to an execution of one task on one cluster node.
Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
In a similar manner, the historical data provided by Ref.~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
size of files that the task sends to all its children.

For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
The same workflow can require different memory capacities and take different times to execute
depending on the size of its input.
We simulate these various runs by obtaining values corresponding to each input size.
For each of the four families, there are five input sizes; we thus run each workflow in five variants corresponding to these inputs.

Not all tasks have historical runtime data stored in the tables.
In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
For two more, around 40\% of the tasks have no historical runtime data stored.
Hence, in the absence of historical data about a task, we give it fixed weights:
an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
These values align with the findings of Ref.~\cite{lotaru} about small tasks.
%
 %   \skug{insert more abt how the runtime system hanfles data reads}

@@

To fully benefit from the historical data, the  {\em default} experimental environment
that we consider is a cluster based on the same six
kinds of real-world machines that were used in the experimental evaluation in Ref.~\cite{lotaru}.
We set the number of each kind of node to $12$, yielding 72 processors in total.  % the whole cluster.
%
Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as tuples (name, speed, memory):
($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and
large memory (see Table~\ref{tab:procs}).

\begin{table}[htb]
    \begin{center}
        \begin{tabular}{c|c|cc}
            \toprule
            Processor  %& Amount
            &  CPU speed   & \multicolumn{2}{c}{Memory size (GB)} \\
            name & (GHz) & {\em default} & {\em mem-constrained} \\
            \midrule
            local                    & 4                    & 16     & 1.6 \\
            A1                      & 32                   & 32     & 3.2 \\
            A2                      & 6                    & 64    & 6.4 \\
            N1                      & 12                   & 16     & 1.6 \\
            N2                      & 8                    & 8      & 0.8\\
            C2                      & 32                   & 192   &  19.2\\
            \bottomrule
        \end{tabular}
    \end{center}
    \caption{Cluster configurations.}
    \label{tab:procs}
\end{table}

We also consider a more constrained setting by varying the cluster configuration. 
The {\em memory-constrained cluster} consists of 72 nodes (12 of each kind) as the default cluster,
but each node has 10 times less memory. Hence,  
the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, 
$A2$ has $6.4$~GB instead, of $64$~GB etc.
The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).
